---
title: "Build and deploy a stroke prediction model using R"
author: "Bhuvaneswari Lakshmanan Pichaimuthu"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# About Data Analysis Report

This RMarkdown file contains the report of the data analysis done for the project on building and deploying a stroke prediction model in R. It contains analysis such as data exploration, summary statistics and building the prediction models. The final report was completed on `r date()`.

**Data Description:**

According to the World Health Organization (WHO) stroke is the 2nd leading cause of death globally, responsible for approximately 11% of total deaths.

This data set is used to predict whether a patient is likely to get stroke based on the input parameters like gender, age, various diseases, and smoking status. Each row in the data provides relevant information about the patient.

# Task One: Import data and data preprocessing

## Load data and install packages

```{r}
# Installing necessary packages for this project
install.packages('tidyverse')
install.packages('skimr')
install.packages('ggplot2')
install.packages("dplyr")
install.packages("tidymodels")
install.packages("Metrics")
```

```{r}

# Loading library files for this project
library(tidyverse)
library(tidymodels)
library(skimr)
library(ggplot2)
library(tidyr)
library(dplyr)

```

```{r}
# Load data into dataframe
hc_stroke_data <- read_csv("healthcare-dataset-stroke-data.csv")

```

## Describe and explore the data

```{r}
colnames(hc_stroke_data)
```

```{r}
# Change column names to be readable
hc_stroke_data <- hc_stroke_data %>% rename("residence_type" = "Residence_type")
```

```{r}
nrow(hc_stroke_data)
n_distinct(hc_stroke_data$id)
is.null(hc_stroke_data)
glimpse(hc_stroke_data)
summary(hc_stroke_data)
head(hc_stroke_data)
```

```{r}
hc_stroke_data <- hc_stroke_data %>% mutate_if(is.character, as.factor)
```

```{r}
summary(hc_stroke_data)
```

```{r}
# From summary above, we found columns "bmi" has N/A values and we have to replace these values to NA values
hc_stroke_data[hc_stroke_data[["bmi"]] == "N/A", "bmi"] <- NA
sum(is.na(hc_stroke_data$bmi))
```


```{r}
# Replace NA values with 0 as bmi as to be numerical
hc_stroke_data <- hc_stroke_data %>%
  mutate(bmi = fct_explicit_na(bmi, "0"))
```

```{r}
# Now change datatype of column "bmi"
hc_stroke_data$bmi <- as.numeric(as.character(hc_stroke_data$bmi))
hc_stroke_data %>% summary()
```

```{r}
# Check for duplicate rows
sum(duplicated(hc_stroke_data))
```

```{r}
# check for empty values
colSums(hc_stroke_data=="")
```

```{r}
# Remoce id from the dataset

hc_stroke_data <- hc_stroke_data[,-1]

```


```{r}
# Remove 1 observation "Other" in gender variable

hc_stroke_data <- hc_stroke_data %>% filter(gender!="Other")

```

```{r}
# Identify Categorial and Numerical variables

categorical_cols <- c("gender","ever_married","work_type","residence_type","smoking_status")
numerical_cols <- c("age","hypertension","heart_disease","avg_glucose_level", "bmi")

```


```{r}
# Plot Gender Vs Stroke
hc_stroke_data %>% aggregate(stroke~gender, FUN=sum) %>% ggplot(aes(x = gender, y=stroke, color=gender, label=stroke)) + geom_text()
```
From the above result, number of females getting stroke is higher than the male.


## Analysis on Numerical variables(Internal Factors) for stroke cases


```{r}
# Plot Age Vs Stroke
hc_stroke_data %>% aggregate(age ~ stroke+gender, FUN=mean) %>% ggplot(aes(x = stroke, y = age, color=gender, label=age)) +
  geom_text()
```

From the above result, stroke cases with average age identified in Female is 67 and in Male is 68.5


```{r}
# Plot Hypertension Vs Stroke
hc_stroke_data %>% aggregate(stroke ~ hypertension+gender, FUN=sum) %>% ggplot(aes(x = hypertension, y = stroke, color=gender, label=stroke)) +
  geom_text()
```

From the above result, hypertension is not a key factor for stroke cases.

```{r}
# Plot Heart Disease Vs Stroke
hc_stroke_data %>% aggregate(stroke ~ heart_disease+gender, FUN=sum) %>% ggplot(aes(x = heart_disease, y = stroke, color=gender, label=stroke)) +
  geom_text()
```
From the above result, heart disease is not a key factor 


```{r}
# Plot Average Glucose Level Vs Stroke
hc_stroke_data %>% aggregate(avg_glucose_level ~ stroke+gender, FUN=mean) %>% ggplot(aes(x=stroke, y=avg_glucose_level, color=gender, label=avg_glucose_level)) +
  geom_text(position=position_dodge(width=0.9))
```


From the above result, average glucose level identified in both genders 143 in Male and 124 in Female.


```{r}
# Plot BMI Vs Stroke
hc_stroke_data %>% aggregate(bmi ~ stroke+gender, FUN=mean) %>% ggplot(aes(x=stroke, y=bmi, color=gender, label=bmi)) +
  geom_text(position=position_dodge(width=0.5))
```


From the above result, Stroke cases with average BMI identified in Female is 25.71 and in Male is 25.39


## Analysis on the Categorical variables(External Factors) for stroke cases

```{r}
# Plot Married Vs Stroke
hc_stroke_data %>% aggregate(stroke ~ ever_married+gender, FUN=sum) %>% ggplot(aes(x=ever_married, y=stroke, color=gender, label=stroke)) + geom_text()
```

Stroke cases seems to be appeared high in married people.

```{r}
# Plot Residence Type Vs Stroke
hc_stroke_data %>% aggregate(stroke ~ Residence_type+gender, FUN=sum) %>% ggplot(aes(x=Residence_type, y=stroke, color=gender, label=stroke)) + geom_text()
```

From the above result, Urban has more stroke cases and cases in Female is higher than Male.


```{r}
# Plot Work type Vs Stroke
hc_stroke_data %>% aggregate(stroke ~ work_type+gender, FUN=sum) %>% ggplot(aes(x=work_type, y=stroke, color=gender, label=stroke)) + geom_text()
```
People with private jobs has high possibilities in getting stroke.


```{r}
# Plot Smoking Vs Stroke
hc_stroke_data %>% aggregate(stroke ~ smoking_status+gender, FUN=sum) %>% ggplot(aes(x=smoking_status, y=stroke, color=gender, label=stroke)) + geom_text()
```
People who formerly smoked has high chances in getting stroke.



```{r}
install.packages("GGally")
library(GGally)
```


## Correlate Stroke with Age, Hypertension, Heart Disease, Average Glucose Level, BMI (Numerical Variables)

```{r}
ggcorr(hc_stroke_data, title="Correlations Between Stroke and (Age, Hypertension, Heart Disease, Average Glucose Level, BMI)")
```


## Check for outliers

From the correlation matrix above, the only outlier is BMI.

# Prepare Data for prediction models

"stroke" variable is set for prediction. 

Remove stroke variable from the data set and split into train and test data sets and predict using Random Forest model.


```{r}
# Set stroke as factor

hc_stroke_data['stroke'] <- lapply(hc_stroke_data['stroke'], factor)

```





# Task Two: Build prediction models

```{r}

```

# Task Three: Evaluate and select prediction models

```{r}

```

# Task Four: Deploy the prediction model

```{r}

```

# Task Five: Findings and Conclusions
